



Software-Defined Networking (SDN) is an emerging technology that offers the required flexibility and interfaces to effectively implement different network monitoring tasks. In fact, an SDN enabler, such as OpenFlow, nicely separates the measurement data plane and control plane functions, and provides a capability to control/re-program the internal configurations of switches in dynamic environments. Consequently, SDN allows for more complex network monitoring and management applications \cite{IF14iSTAMP:2014} \cite{YLiao:2011} \cite{MYu:2011} \cite{MYu:2013}. %  in an adaptive and efficient way

Software-Defined Networking (SDN) is an emerging technology that offers the required flexibility and interfaces to effectively implement more complex network monitoring and management applications. In fact, an SDN enabler, such as OpenFlow, nicely separates the measurement data plane and control plane functions, and provides a capability to control/re-program the internal configurations of switches in dynamic environments. Consequently, SDN allows for more complex network monitoring and management applications \cite{IF14iSTAMP:2014} \cite{YLiao:2011} \cite{MYu:2011} \cite{MYu:2013}. %  in an adaptive and efficient way


Software-Defined Networking (SDN), along with its enabler such as OpenFlow, is an emerging technology that nicely separates the measurement data plane and control plane functions, and provides a capability to control/re-program the internal configurations of switches in dynamic environments. Consequently, SDN allows to effectively and efficiently implement more complex network monitoring applications, including both passive and active network measurements, without the need of customization \cite{IF14iSTAMP:2014} \cite{YLiao:2011} \cite{MYu:2011} \cite{MYu:2013}. For passive network measurement, the most SDN based works related to traffic engineering and network security applications where the main goal is focused on network traffic measurement or identifying aspects of the network traffic such the presence of Heavy Hitters (HH) and Hierarchical Heavy Hitters (HHH). In \cite{MYu:2011}, the authors propose a reconfigurable measurement architecture for hierarchical heavy hitter detection, and then in \cite{MYu:2013}, they propose a re-programmable structure (called OpenSketch) where a variety of sketches for different measurement tasks can be defined and installed by the operator. In \cite{Tootoonchian:2010}, OpenTM estimates a traffic matrix by keeping track of statistics for each flow. Recently, in \cite{IF14iSTAMP:2014}, we have proposed an intelligent SDN based traffic measurement framework (called iSTAMP) with the ability of adaptive and accurate fine-grained flow estimation. In addition, for active network measurement under SDN paradigm, the very recent work \cite{Adrichen:2014} establishes a general framework where accurate measurements of flow throughput, packet loss and delay can be obtained.





 \cite{IF14iSTAMP:2014} \cite{YLiao:2011} \cite{MYu:2011} \cite{MYu:2013}.

that offers the required flexibility and interfaces to effectively implement more complex network monitoring and management applications. In fact, an SDN enabler, such as OpenFlow, nicely separates the measurement data plane and control plane functions, and provides a capability to control/re-program the internal configurations of switches in dynamic environments. Consequently, SDN allows for more complex network monitoring and management applications \cite{IF14iSTAMP:2014} \cite{YLiao:2011} \cite{MYu:2011} \cite{MYu:2013}. %  in an adaptive and efficient way



The flexibility provided by SDN and its enabler OpenFlow has been used to implement most of the passive and active network monitoring tasks in different applications without the need of customization. For passive network measurement, the most SDN based works related to traffic engineering and network security applications where the main goal is focused on network traffic measurement or identifying aspects of the network traffic such the presence of Heavy Hitters and Hierarchical heavy hitters. In \cite{MYu:2011}, the authors propose a reconfigurable measurement architecture for hierarchical heavy hitter detection, and then, \cite{MYu:2013} they propose a re-programmable structure (called OpenSketch) where a variety of sketches for different measurement tasks can be defined and installed by the operator. In \cite{Tootoonchian:2010}, OpenTM estimates a traffic matrix by keeping track of statistics for each flow. Recently, in \cite{IF14iSTAMP:2014} we have proposed an intelligent SDN based traffic measurement framework (called iSTAMP) with the ability of adaptive and accurate fine-grained flow estimation. For active network measurement under SDN paradigm, the very recent work \cite{Adrichen:2014} establishes a general framework where accurate measurements of flow throughput, packet loss and delay can be obtained.
