\begin{abstract}
A key requirement for network management is accurate and reliable network monitoring where critical information about internal characteristics or states of the network(s) must be obtained. In today's large-scale networks, this is a challenging task due to the hard constraints of network measurement resources. In this paper, a new framework (called SNIPER) is proposed where we use the flexibility provided by Software-Defined Networking (SDN) to design the optimal observation or measurement matrix which leads to the best achievable estimation accuracy using Matrix Completion (MC) techniques. Here, to cope with the inherent complexity of the process of designing large-scale optimal observation matrices, we use the well known Evolutionary Optimization Algorithms (EOA) which directly target the ultimate estimation accuracy as the optimization objective function. We evaluate the performance of SNIPER using both synthetic and real network measurement traces from different network topologies and by considering two main applications including network traffic and delay estimations. Our results show that SNIPER is a flexible and efficient framework that can be used for a variety of network performance measurements under hard constraints of network measurement resources. For example, by measuring 8.8\% of per-flow path delays in Harvard network, congested paths can be detected by probability 0.94. To demonstrate the feasibility and effectiveness of our framework, we have implemented a prototype of SNIPER in Mininet.
\end{abstract}